%!TEX root = std.tex
\chapter{Foreword}

ISO (the International Organization for Standardization) and IEC (the
International Electrotechnical Commission) form the specialized system for
worldwide standardization. National bodies that are members of ISO or IEC
participate in the development of International Standards through technical
committees established by the respective organization to deal with particular
fields of technical activity. ISO and IEC technical committees collaborate in
fields of mutual interest. Other international organizations, governmental and
non-governmental, in liaison with ISO and IEC, also take part in the work. In
the field of information technology, ISO and IEC have established a joint
technical committee, ISO/IEC JTC 1.

The procedures used to develop this document and those intended for its further
maintenance are described in the ISO/IEC Directives, Part 1. In particular the
different approval criteria needed for the different types of document should
be noted. This document was drafted in accordance with the editorial rules of
the ISO/IEC Directives, Part 2
(see~\href{http://www.iso.org/directives}{www.iso.org/directives}).

Attention is drawn to the possibility that some of the elements of this
document may be the subject of patent rights. ISO and IEC shall not be held
responsible for identifying any or all such patent rights.  Details of any
patent rights identified during the development of the document will be in the
Introduction and/or on the ISO list of patent declarations received
(see~\href{http://www.iso.org/patents}{www.iso.org/patents}).

Any trade name used in this document is information given for the convenience
of users and does not constitute an endorsement.

For an explanation on the meaning of ISO specific terms and expressions related
to conformity assessment, as well as information about ISO's adherence to the
WTO principles in the Technical Barriers to Trade (TBT) see the following URL:
\href{http://www.iso.org/iso/home/standards_development/resources-for-technical-work/foreword.htm}{Foreword - Supplementary information}

The committee responsible for this document is
ISO/IEC JTC 1, \textit{Information technology},
SC 22, \textit{Programming languages, their environments and system software interfaces}

This fifth edition cancels and replaces the fourth edition
(ISO/IEC 14882:2014), of which it constitutes a minor revision.
